---
title: "My Portfolio"
author: "Chris"
date: "2024-02-21"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    css: stylesheets/style.css
---

```{r setup, include=FALSE}
remotes::install_github('jaburgoyne/compmus')

library(tidyverse)
library(spotifyr)
library(compmus)
library(cowplot)
library(plotly)
```

```{r corpus, include=FALSE}
lettberg <- readRDS(file = "data/playlists/lettberg-playlist.RDS")
ashkenazy <- readRDS(file = "data/playlists/ashkenazy-playlist.RDS")
alexeev <- readRDS(file = "data/playlists/alexeev-playlist.RDS")
zhukov <- readRDS(file = "data/playlists/zhukov-playlist.RDS")

pianists <- list(lettberg, ashkenazy, alexeev, zhukov)
all_recordings <- bind_rows(pianists) %>%
    mutate(period = replace(sonata, sonata < 11, "Late")) %>%
    mutate(period = replace(period, sonata < 6, "Middle")) %>%
    mutate(period = replace(period, sonata < 4, "Early"))
```

```{r functions_chromagram, include=FALSE}
chroma_preprocessing <- function(sonata) {
  sonata <- sonata |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)
}

chromagram <- function(sonata, chroma) {
  if (chroma) {
    plot <- ggplot(sonata,
         aes(x = start + duration / 2,
            width = duration,
            y = pitch_class,
            fill = value
            )
        )
  } else {
    plot <- ggplot(sonata,
         aes(x = start + duration / 2,
            width = duration,
            y = basis,
            fill = value
            )
        )
  }
  
  plot +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude") +
  theme_classic() +
  scale_fill_viridis_c() +
  scale_x_continuous(breaks=seq(0,2000,60))
}
```

### It is difficult to gather much information out of timbre

```{r  self_similarity, include=FALSE}
lettberg_1_1 <- readRDS(file = "data/pieces/lettberg/lettberg_1_1.RDS")
lettberg_2_1 <- readRDS(file = "data/pieces/lettberg/lettberg_2_1.RDS")
lettberg_3_1 <- readRDS(file = "data/pieces/lettberg/lettberg_3_1.RDS")
lettberg_4_1 <- readRDS(file = "data/pieces/lettberg/lettberg_4_1.RDS")
lettberg_5_1 <- readRDS(file = "data/pieces/lettberg/lettberg_5_1.RDS")
lettberg_6_1 <- readRDS(file = "data/pieces/lettberg/lettberg_6_1.RDS")
lettberg_7_1 <- readRDS(file = "data/pieces/lettberg/lettberg_7_1.RDS")
lettberg_8_1 <- readRDS(file = "data/pieces/lettberg/lettberg_8_1.RDS")
lettberg_9_1 <- readRDS(file = "data/pieces/lettberg/lettberg_9_1.RDS")
lettberg_10_1 <- readRDS(file = "data/pieces/lettberg/lettberg_10_1.RDS")
```

```{r functions_timbre_selfsim, include=FALSE}
chroma_and_timbre <- function(sonata) {
  sonata <- sonata |>
  compmus_align(bars, segments) |>                     # Change `bars`
  select(bars) |>                                      #   in all three
  unnest(bars) |>                                      #   of these lines.
  mutate(
    pitches =
      map(segments,
        compmus_summarise, pitches,
        method = "rms", norm = "euclidean"              # Change summary & norm.
      )
  ) |>
  mutate(
    timbre =
      map(segments,
        compmus_summarise, timbre,
        method = "rms", norm = "euclidean"             # Change summary & norm.
      )
  )
}

self_similarity <- function(sonata, chroma) {
  if (chroma) {
    plot <- sonata |>
      compmus_self_similarity(pitches, "cosine") 
  } else {
    plot <- sonata |>
      compmus_self_similarity(timbre, "cosine") 
  }
  
  plot |> 
    ggplot(
      aes(
        x = xstart + xduration / 2,
        width = xduration,
        y = ystart + yduration / 2,
        height = yduration,
        fill = d
      )
    ) +
    geom_tile() +
    coord_fixed() +
    scale_fill_viridis_c(guide = "none") +
    theme_classic() +
    labs(x = "", y = "") +
    scale_x_continuous(breaks=seq(0,2000,60)) +
    scale_y_continuous(breaks=seq(0,2000,60)) +
    theme(axis.text.x = element_text(angle=-45))
}
```

```{r}
timbre_analysis <- list(lettberg_1_1,
                        lettberg_8_1)

timbre_analysis <- lapply(timbre_analysis, chroma_and_timbre)
```

```{r}
sonata_1 <- timbre_analysis[[1]]
sonata_8 <- timbre_analysis[[2]]


cep_1 <- sonata_1 |>
  compmus_gather_timbre() |>
  chromagram(chroma = FALSE) +
    geom_vline(xintercept = 151, colour = "red") +
    geom_vline(xintercept = 299, colour = "red") +
    labs(title = "Cepstrogram of the 1st sonata", subtitle = "Maria Lettberg")


cep_8 <- sonata_8 |>
  compmus_gather_timbre() |>
  chromagram(chroma = FALSE) +
  labs(title = "Cepstrogram of the 8th sonata", subtitle = "Maria Lettberg") +
  geom_vline(xintercept = 140, colour = "red")


plot_grid(plotlist = list(cep_1, cep_8), ncol = 1)
```

***
Similarly to the chromagrams, spotify also allows us to analyse the timbre content of the pieces, referred to as Cepstrograms.
However, solo piano works don't always lend themselves to this type of analysis, since you are generally limited in the kinds of timbres
you can create with just a piano, unless you start employing extended techniques like in [George Crumb's](https://youtu.be/pIEUS0-DDcs) case for instance. Regardless, there are instances where a pieces's eccentricities are reflected in its cepstrogram.

Consider a sonata's structure for example. One defining aspect of the sonata is how its first movement is composed in sonata form. Although like most musical conventions, composers have been straying from this practice over time, including Scriabin. Despite that, his first sonata does display a strong semblance of sonata form. When looking at the cepstrogram, you can make out how the exposition section is repeated once in the sonata's beginning.

Structural details like this are hard to find in his other sonatas, especially when only looking at pitch and timbre. It is quite tough to define an overarching structure in his 8th sonata for instance. Nevertheless, you can clearly see the first 2 minutes being distinct from the rest of the piece. Upon giving this sonata a listen, you'll notice that this section is noticeably unnerving yet tranquil.

### The structure of Scriabin's sonatas is not always easy to define

```{r}
timbre_analysis <- list(lettberg_2_1,
                        lettberg_7_1)

timbre_analysis <- lapply(timbre_analysis, chroma_and_timbre)

sonata_2 <- timbre_analysis[[1]]
sonata_7 <- timbre_analysis[[2]]
```

```{r}
selfsim_chroma_1 <- self_similarity(sonata_1, chroma=TRUE) +
  labs(title = "Pitch | Sonata No. 1", subtitle = "Maria Lettberg")
selfsim_timbre_1 <- self_similarity(sonata_1, chroma=FALSE) +
  labs(title = "Timbre | Sonata No. 1", subtitle = "Maria Lettberg")

plot_grid(plotlist = list(selfsim_chroma_1, selfsim_timbre_1)) + labs(title = "Self similarity matrices")
```

***
These plots display the pitch and timbre-based self-similarity within Scriabin's first sonata. While the Timbre matrix does not seem to divulge much, since the overall timbre of the piece does not vary much, you can still make out how there's one bar after around 150 seconds where there's a large pause. Other than that, the plot does not tell us much.

The pitch-based self-similarity matrix on the other hand does clearly display the repetition of the sonata's exposition. As seen in how there's a dark blue line parallel to the diagonal of the matrix, indicating that two different sections of the piece overlap. The development and recapitulation are a bit harder to make out. However if you look closely, you can just barely see a very short dark blue diagonal line starting at 420 seconds on the x axis. This line is very short since the recapitulation almost immediately modulates to a different key.

### Self-similarity matrices help very little with the later sonatas

```{r}
selfsim_chroma_7 <- self_similarity(sonata_7, chroma=TRUE) +
  labs(title = "Pitch | Sonata No. 7", subtitle = "Maria Lettberg")
selfsim_timbre_7 <- self_similarity(sonata_7, chroma=FALSE) +
  labs(title = "Timbre | Sonata No. 7", subtitle = "Maria Lettberg")

plot_grid(plotlist = list(selfsim_chroma_7, selfsim_timbre_7))
```

***
As Scriabin started breaking away from conventional late-romanticism, he also began applying structure to his pieces in a different way.

### The style of Scriabin 

There are many composers who have made a name for themselves when it comes to their contribution to the piano repertoire. One such composer is Alexander Scriabin. However, what seperates him from other composers of similar regard, is the way in which his musical language has developed throughout his life. The influence of Frederic Chopin and other romantic composers can clearly be discerned at the start of his carreer. In spite of that, his music would increasingly veer away from traditional harmony and tonality over the turn of the 20th century. Much like Arnold Schoenberg, he had developed a unique, modernist style of music that would influence many other 20th century composers.

One of my goals is to examine wether Scriabin's changing style could be accurately represented. This development is perhaps best reflected in Scriabin's 10 piano sonatas, which span the majority of his composing carreer. My corpus therefore consists of recordings of these sonatas by four pianists, namely Vladimir Ashkenazy, Maria Lettberg, Igor Zhukov and Dmitri Alexeev. These pieces, much like his other works, leave the performer with room for their own interpretation. So It would perhaps also be interesting to see in what capacity spotify could measure the difference between performers.

When considering tracks of particular interest, the 4th and 5th piano sonatas come to mind. The former being one of his first pieces to display a concrete change in style, and the later being one of the first stepts into truly new territory. Yet despite their differences they do both have a distinct touch of lightness that is in my opinion more difficult to find in his other sonatas. A further look into these pieces may yield intriguing results.

### Comparing the energy and valence for all sonatas

```{r Energy_valence}
all_recordings_mean_energy_valence <- all_recordings %>%
  group_by(pianist, period, sonata) %>%
  summarize(mean_energy = mean(energy),
            mean_valence = mean(valence),
            period = factor(period, levels=c("Early", "Middle", "Late")))

zhukov_mean_energy_valence <- all_recordings_mean_energy_valence %>%
  filter(pianist == "Igor Zhukov")

cols <- c("Early" = "#c2df23", "Middle" = "#1fa187", "Late" = "#440154")

zhukov_plot <- ggplot(zhukov_mean_energy_valence, aes(x = mean_energy, y = mean_valence, color= period)) +
  annotate("rect", xmin = 0.065, xmax = 0.15, ymin = 0.028, ymax = 0.038, alpha = .2) +
  geom_point() +
  scale_y_log10() +
  scale_x_log10() +
  # scale_color_grey(start = 0.75, end = 0) +
  labs(x="Energy", y="Valence", color="Period", title="Energy and Valence for Scriabin's 10 sonatas",
       subtitle="Igor Zhukov | The late sonatas have a distinctly low valence level.") +
  annotate(
    "text",
    x = 0.1, y = 0.04,
    label = "Scriabin's late sonatas",
    vjust = 1, size = 5, color = "grey40"
  ) +
  scale_color_manual(values = cols)

ggplotly(zhukov_plot)
```

***

This plot displays the valence and energy levels of the sonatas performed by Igor Zhukov, just as an example. If you focus on the valence feature, valence being a measure of positiveness, you might notice that while the differences accross sonatas are small, all the late sonatas seem to lay at a distinct valence level compared to the rest.

Those familiar with the music in question might not be suprised by this. These pieces are very representative of Scriabin's changing musical style, despite the fact that they are individually still unique. 


### The late sonatas contain similar valence levels accros all performers

```{r Energy_Valence_all}
energy_valence_plot <- ggplot(all_recordings_mean_energy_valence, aes(x = mean_energy, y = mean_valence, color= period)) +
  geom_point() +
  facet_wrap(~ pianist) +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_manual(values = cols) +
  labs(x="Energy", y="Valence", color="Period", title="Energy and Valence for Scriabin's 10 sonatas",
       subtitle="The late sonatas have a distinctly low valence level.")

ggplotly(energy_valence_plot)
```

***

here you can see the previous results extend very well to other pianists and their recordings. While the energy and valence levels of the early and middle period of Scriabin's works vary quite a bit, the late sonatas all fall on the same range of valence, even across different performers.

At the same time however, the pieces show somewhat of a variance in energy levels across performers. This could perhaps reflect the natural room for interpretation of Scriabin's works, albeit this is hard to say from just this data.

### How clearly does a chromagram show a sonata's key?

```{r get_sonatas_chromagram, include=FALSE}
ashkenazy_4_1 <- readRDS(file = "data/pieces/ashkenazy/ashkenazy_4_1.RDS")
ashkenazy_4_2 <- readRDS(file = "data/pieces/ashkenazy/ashkenazy_4_2.RDS")
ashkenazy_5 <- readRDS(file = "data/pieces/ashkenazy/ashkenazy_5_1.RDS")

chroma_comparison <- list(ashkenazy_4_1,
                          ashkenazy_4_2,
                          ashkenazy_5)

chroma_comparison <- lapply(chroma_comparison, chroma_preprocessing)
```

```{r process_4_5, include=FALSE}
ashkenazy_4_1_chroma <- chroma_comparison[[1]] %>%
  mutate(movement = 1)

ashkenazy_4_2_chroma <- chroma_comparison[[2]] %>%
  mutate(movement = 2,
         start = start + tail(ashkenazy_4_1_chroma, n=1)$start)

ashkenazy_4_pitches <- bind_rows(list(ashkenazy_4_1_chroma, ashkenazy_4_2_chroma)) %>%
    mutate(sonata = "Sonata 4")

ashkenazy_5_pitches <- chroma_comparison[[3]] %>%
  mutate(sonata = "Sonata 5")
```

```{r Chromagram_4_5}
ashkenazy_4_chroma <- ashkenazy_4_pitches |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma()

ashkenazy_5_chroma <- ashkenazy_5_pitches |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma()


chromagram_ashk_4 <- chromagram(ashkenazy_4_chroma, TRUE) +
    labs(title = "Chromagram of the 4th sonata", subtitle = "Vladimir Ashkenazy") +
    geom_vline(xintercept = head(ashkenazy_4_2_chroma, n=1)$start, colour = "red")

chromagram_ashk_5 <- chromagram(ashkenazy_5_chroma, TRUE) +
    labs(title = "Chromagram of the 5th sonata", subtitle = "Vladimir Ashkenazy")

chromagram_ashk_4_5 <- list(chromagram_ashk_4, chromagram_ashk_5)

plot_grid(plotlist = chromagram_ashk_4_5, ncol = 1)
```

***
Throughout Scriabin's carreer as a composer, you can quite clearly hear how his music starts getting more and more tonally ambiguous. Even to the point where it might not make much sense to discuss it in terms of tonality.

Take the following chromagrams of the 4th and 5th sonata for example. It is not far fetched to say that Scriabin's music has changed the most between these pieces. As a result, you can, albeit without the most confidence, observe how this piece is in F#. The 5th sonata on the other hand does not display display a clear tonic like the previous one does, which makes sense upon giving it a listen.

### From the 5th sonata, each sonata's key starts getting more ambiguous

```{r load_ashkenazy, include=FALSE, cache=TRUE}
ashkenazy_2_1 <-
  get_tidy_audio_analysis("2vNz6dat4cjEJol2AoqIIt") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

ashkenazy_2_2 <-
  get_tidy_audio_analysis("6pHp6L7zlZDO5LClxp3pYN") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches) %>%
  mutate(start = start + tail(ashkenazy_2_1, n=1)$start)

ashkenazy_2 <- bind_rows(list(ashkenazy_2_1, ashkenazy_2_2)) %>%
  mutate(sonata = "Sonata 2") |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
    ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Sonata 2") +
  theme_minimal() +
  scale_fill_viridis_c()

ashkenazy_4_chroma <- ashkenazy_4_chroma |>
  ggplot(aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
    ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Sonata 4") +
  theme_minimal() +
  scale_fill_viridis_c()

ashkenazy_6 <-
  get_tidy_audio_analysis("0NaVN5lDbMkmSwxSiAXMVc") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches) %>%
  mutate(sonata = "Sonata 6") |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
    ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Sonata 6") +
  theme_minimal() +
  scale_fill_viridis_c()

ashkenazy_8 <-
  get_tidy_audio_analysis("5RKOyVBXa05wImhToOBPcq") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches) %>%
  mutate(sonata = "Sonata 8") |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
    ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Sonata 8") +
  theme_minimal() +
  scale_fill_viridis_c()

ashkenazy_10 <-
  get_tidy_audio_analysis("27H4M1wOVah2qqHq5xbkeM") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches) %>%
  mutate(sonata = "Sonata 10") |>
  mutate(pitches = map(pitches, compmus_normalise, "euclidean")) |>
  compmus_gather_chroma() |>
  ggplot(aes(
      x = start + duration / 2,
      width = duration,
      y = pitch_class,
      fill = value
    )
    ) +
  geom_tile() +
  labs(x = "Time (s)", y = NULL, fill = "Magnitude", title = "Sonata 10") +
  theme_minimal() +
  scale_fill_viridis_c()

chroma_ashkenazy <- list(ashkenazy_2, ashkenazy_4_chroma, ashkenazy_6, ashkenazy_8, ashkenazy_10)

```

```{r Chromagram_ashkenazy, fig.height=16, fig.width=10}
plot_grid(plotlist = chroma_ashkenazy, ncol = 1)
```

***
Here are a couple more chromagrams. Let's look at a couple more pieces, once again performed by Vladimir Ashkenazy just for an example.

<br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br> <br>

### Aligning different performances

```{r DTW, cache=TRUE}
lettberg_5 <-
  get_tidy_audio_analysis("0bCTqxYSyYP7OawV9cam1V") |>
  select(segments) |>
  unnest(segments) |>
  select(start, duration, pitches)

DTW_plot <- compmus_long_distance(
  ashkenazy_5_pitches |> mutate(pitches = map(pitches, compmus_normalise, "chebyshev")),
  lettberg_5 |> mutate(pitches = map(pitches, compmus_normalise, "chebyshev")),
  feature = pitches,
  method = "euclidean"
) |>
  ggplot(
    aes(
      x = xstart + xduration / 2,
      width = xduration,
      y = ystart + yduration / 2,
      height = yduration,
      fill = d
    )
  ) +
  geom_tile() +
  coord_equal() +
  labs(x = "Vladimir Ashkenazy", y = "Maria Lettberg", title = "Sonata 5") +
  theme_minimal() +
  scale_fill_viridis_c(guide = NULL)

DTW_plot
```

***
Text.

